Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-03-24 bbbq_article.tex
Moved 'Prediction software written' to SupMat
Moved Frans' text to Methods

For MHC-I, 9-mers were used, as this is the length used in our earlier
study (\cite{bianchi2017}). For MHC-II, 14-mers were used, 
as these are the most frequently occurring
epitope length (\cite{bergseng2015different}).
We used a human (UniProt ID UP000005640\_9606), 
viral (SARS-CoV-2, UniProt ID UP000464024) 
and bacterial (Mycobacterium tuberculosis (MTb), UniProt ID UP000001584) 
reference proteome. The pathogens were chosen based 
on their impact on the human 
population \richel{ref}.
We used TMHMM  \richel{ref} to predict the topology 
of the proteins within the proteome.
To predict the affinity of an epitope to a certain MHC molecule,
we used \verb;EpitopePrediction; (\cite{bianchi2017}) for MHC-I 
and \verb;MHCnuggets; \richel{ref} for MHC-II.
The 13 MHC-I haplotypes used in this study are the same as 
used in our previous study (\cite{bianchi2017}).
The MHC-II haplotypes used additionally were selected 
to occur with a phenotypic frequency of at least 14\% in
the human population (\cite{greenbaum2011functional}),
resulting in 21 haplotypes.

@@

[RJCB: I suggest to move this section to the Supplementary Materials]
\frans{I agree move to supp}

@@

To confirm that peptides stemming from TMHs were presented in MHC-I and MHC-II,
peptide elution studies were reanalyzed.
Table \ref{tab:elution} shows the percentage of epitopes derived 
from a TMH
found in MHC-I and MHC-II elution 
studies (\cite{schellens2015comprehensive} and \cite{bergseng2015different} respectively),
for the two topology prediction tools TMHMM and 
PureseqTM (\cite{krogh2001predicting} and \cite{wang2019efficient} respectively). 
Regardless of the prediction, 
at least 100 epitopes were predicted to be derived from a TMH. 
From these findings, it is robustly predicted that
epitopes derived from TMHs are presented by B cells [For Frans: please check
that the article by Bergseng, 2015, 'Different binding motifs of the celiac 
disease-associated HLA molecules DQ2. 5, DQ2. 2, and DQ7. 5 revealed by 
relative quantitative proteomics of endogenous peptide repertoires'
is appropriate. As far as I can see, it is, but I may overlook something. Richel].

@@

We next wondered if this over-representation would also generalize to MHC-II. \frans{ you say MHC-II but this is very general I think you have to explain the choice for these haplotypes.}
Unlike MHC-I where only one variable chain determines the landscape of the epitope binding pocket allowing for the classification of 5 HLA-A and 8 HLA-B superhaplotypes with .... variation in peptide binding repertoire, the MHC-II peptide binding groove is determined by both the alpha and beta chain. For HLA DQ both the alpha and beta chain is highly polymorphic  ..... \frans{why is HLA-DP not predicted?}
We took the alleles that represent percentage of the occurrence int he overall human population according to ... et al. 
Figure \ref{fig:2} shows the percentages of MHC-II epitopes overlapping 
with TMHs for our human, viral and bacterial proteome.
See the supplementary materials (table \ref{tab:tmh_binders_mhc2}) 
for the TMH and epitope counts.
We show that MHC-II has the same pattern of over-presentation
as MHC-I, yet this over-presentation is more pronounced. \frans{we should explain this does it like shown for mhc-I also linearly correlates with the hydrophopbicity of the MHC-II binding celft?}
Also for MHC-II, we observe that the extent to which TMH-derived
epitopes are presented is similar between the haplotypes, for
a human, bacterial and viral proteome.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}\frans{why comes this section before discussion?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

>>>>

For MHC-I, 9-mers were used, as this is the length used in our earlier
study (\cite{bianchi2017}). For MHC-II, 14-mers were used, 
as these are the most frequently occurring
epitope length (\cite{bergseng2015different}).
We used a human (UniProt ID UP000005640\_9606), 
viral (SARS-CoV-2, UniProt ID UP000464024) 
and bacterial (Mycobacterium tuberculosis (MTb), UniProt ID UP000001584) 
reference proteome. The pathogens were chosen based 
on their impact on the human 
population \richel{ref}.
We used TMHMM  \richel{ref} to predict the topology 
of the proteins within the proteome.
To predict the affinity of an epitope to a certain MHC molecule,
we used \verb;EpitopePrediction; (\cite{bianchi2017}) for MHC-I 
and \verb;MHCnuggets; \richel{ref} for MHC-II.
The 13 MHC-I haplotypes used in this study are the same as 
used in our previous study (\cite{bianchi2017}).
The MHC-II haplotypes used additionally were selected 
to occur with a phenotypic frequency of at least 14\% in
the human population (\cite{greenbaum2011functional}),
resulting in 21 haplotypes.
\richel{This is Frans' text: 
Unlike MHC-I where only one variable chain determines the landscape of the epitope binding pocket allowing for the classification of 5 HLA-A and 8 HLA-B superhaplotypes with .... variation in peptide binding repertoire, the MHC-II peptide binding groove is determined by both the alpha and beta chain. For HLA DQ both the alpha and beta chain is highly polymorphic  ..... \frans{why is HLA-DP not predicted?}
We took the alleles that represent percentage of the occurrence int he overall human population according to ... et al. 
}

@@

@@

To confirm that peptides stemming from TMHs were presented in MHC-I and MHC-II,
peptide elution studies were reanalyzed.
Table \ref{tab:elution} shows the percentage of epitopes derived 
from a TMH
found in MHC-I and MHC-II elution 
studies (\cite{schellens2015comprehensive} and \cite{bergseng2015different} respectively),
for the two topology prediction tools TMHMM and 
PureseqTM (\cite{krogh2001predicting} and \cite{wang2019efficient} respectively). 
Regardless of the prediction, 
at least 100 epitopes were predicted to be derived from a TMH. 
From these findings, it is robustly predicted that
epitopes derived from TMHs are presented by B cells.

@@

We next wondered if this over-representation would also generalize to MHC-II. 
Figure \ref{fig:2} shows the percentages of MHC-II epitopes overlapping 
with TMHs for our human, viral and bacterial proteome.
See the supplementary materials (table \ref{tab:tmh_binders_mhc2}) 
for the TMH and epitope counts.
We show that MHC-II has the same pattern of over-presentation
as MHC-I, yet this over-presentation is more pronounced. 
\frans{
  we should explain this does it like shown for mhc-I also 
  linearly correlates with the hydrophopbicity 
  of the MHC-II binding cleft?
}
Also for MHC-II, we observe that the extent to which TMH-derived
epitopes are presented is similar between the haplotypes, for
a human, bacterial and viral proteome.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
